\section{Obtaining Test Data}
\label{sec:test_data} 

\subsection{SDDL}

Unfortunately, no ASTERIX test data for the SDDL decoder is available. If you could provide such data, please contact the author. \\

For a short usage guide how to obtain and use SDDL, please refer to \nameref{sec:sddl}.

\subsection{OpenSky Network}

OpenSky ADS-B JSON data can be downloaded from \href{https://opensky-network.org/datasets/states/}{here}, please refer to \href{https://opensky-network.org/}{OpenSky} about licensing details. \\

If a state file is dowloaded, please make sure it is a JSON dataset, like 'states\_2018-12-10-12.json.tar'. After download, extract the GZIP file from it, e.g. 'states\_2018-12-10-11.json.gz'. Only such files can be imported.

\subsection{ADS-B Exchange}

Until recently, data from \href{https://www.adsbexchange.com/}{ADS-B Exchange} was made available for free. Now a small fee must be paid, for hosting and bandwidth costs. Please refer to href{https://www.adsbexchange.com/data/}{historical data} for details.

Under the previous licene, as test datasets, imported dataset from ADSBexchange have been made available with the providers permission. For information about ADS-B exchange please refer to section \nameref{sec:adsbex}. \\

Two datasets can be downloaded:

\begin{table}[H]
  \center
  \begin{tabular}{ | l | l | l |}
    \hline
    \textbf{Dataset} & Size & \textbf{Link} \\ \hline
    2017-11-30\_full.db & 2.1GB & \href{https://drive.google.com/open?id=19JgE8kgVG2lodIyI6Vo8ac7aNY7pM1i_}{Link} \\ \hline
    2017-11-30\_europe.db & 800MB & \href{https://drive.google.com/open?id=1s73R9IEq_8KePC96-a2_seXxZ-Ybhisd}{Link} \\
    \hline
  \end{tabular}
  \caption{Test datasets}
\end{table}

Please be aware that using the produced data for commercial purposes (without a specific agreement) would violate the ADSB exchange terms \& conditions. Please refer to \url{https://www.adsbexchange.com/legal-and-privacy/} for additional information.

