\chapter{Test Data}
\label{sec:test_data} 

\section{ASTERIX Data}

Since it is assumed that most users are from ANSPs, ASTERIX data from their operational or test system should be available. \\

Unfortunately no user has yet made ASTERIX data available to be shared publicly. If you could provide such data, please contact the author. \\

ASTERIX data can be decoded and tested using either jASTERIX or SDDL.

\subsection{jASTERIX}

For a short usage guide how to obtain and use jASTERIX, please refer to \nameref{sec:jasterix}.

\subsection{SDDL}

For a short usage guide how to obtain and use SDDL, please refer to \nameref{sec:sddl}.

\section{OpenSky Network Data}

OpenSky ADS-B JSON data can be downloaded from \href{https://opensky-network.org/datasets/states/}{here}, please refer to \href{https://opensky-network.org/}{OpenSky} about licensing details. \\

If a state file is dowloaded, please make sure it is a JSON dataset, like 'states\_2018-12-10-12.json.tar'. After download, extract the GZIP file from it, e.g. 'states\_2018-12-10-11.json.gz'. Only such files can be imported.

\section{ADS-B Exchange Data}

Until recently, data from \href{https://www.adsbexchange.com/}{ADS-B Exchange} was made available for free. Now a small fee must be paid, for hosting and bandwidth costs. Please refer to href{https://www.adsbexchange.com/data/}{historical data} for details.

Under the previous licene, as test datasets, imported dataset from ADSBexchange have been made available with the providers permission. For information about ADS-B exchange please refer to section \nameref{sec:adsbex}. \\

Two datasets can be downloaded:

\begin{table}[H]
  \center
  \begin{tabular}{ | l | l | l |}
    \hline
    \textbf{Dataset} & Size & \textbf{Link} \\ \hline
    2017-11-30\_full.db & 2.1GB & \href{https://drive.google.com/open?id=19JgE8kgVG2lodIyI6Vo8ac7aNY7pM1i_}{Link} \\ \hline
    2017-11-30\_europe.db & 800MB & \href{https://drive.google.com/open?id=1s73R9IEq_8KePC96-a2_seXxZ-Ybhisd}{Link} \\
    \hline
  \end{tabular}
  \caption{Test datasets}
\end{table}

Please be aware that using the produced data for commercial purposes (without a specific agreement) would violate the ADSB exchange terms \& conditions. Please refer to \url{https://www.adsbexchange.com/legal-and-privacy/} for additional information.

